---
title: "Step 04: Models - ARIMA and Prophet"
output:
  html_document:
    df_print: paged
---
  
## Setup    
  Load the relevant libraries.
```{r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()

# data manipulation
library("plyr")
library("tidyverse")
library("magrittr")
library("lubridate")

# time series specific packages
library("timetk")
library("zoo")
library("tibbletime")

# modeling
library("fpp2")
library("prophet")

# other
library("rsample")            # rolling samples for validation stats
library("ggplot2")            # viz
library("sweep")              # more easily pull out model statistics
library("yardstick")          # easily calculate accuracy stats
library("doParallel")         # parallel processing

```
  
    
  Session Info.
```{r}

sessionInfo()

```
  
    
  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/Bikeshare_Status_SF/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```


## Modeling

**NOTE: `period_train`, `period_test`, and `skip_span`, are the outputs produced in Step 02 and Step 03**
```{r}

period_train <- # (12 * 24 * 365 * 0.5) # 0.5 years of training data
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "period_train.Rds"
                 )
          )

period_test <- # (12 * 24 * 365 * 0.25) # 0.25 years of testing data
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "period_test.Rds"
                 )
          )

skip_span <- # (12 * 24 * 25) # gives 10 evenly distributed  splits
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "skip_span.Rds"
                 )
          )

```  
  
    
  Creating the datasets used in modeling.
```{r}

lmtd_vars_lag3 <-
  lmtd_vars_lag3 %>% 
  map(~select(.x, station_id:minute.55 ))

# small_test <- small_test[1:2]

message("small_test")
names(small_test)
small_test$`10` %>% class()
small_test$`10` %>% glimpse()

message("lmtd_vars_lag3")
lmtd_vars_lag3$`10` %>% class()
lmtd_vars_lag3$`10` %>% glimpse()


# this odd combination of one-hot encoded variables and "normal" variables is only necessary so that a single dataset can be used for Prophet models with "additional regressors"
combo <-
  pmap(.l = list(a = small_test,
                 b = lmtd_vars_lag3
                 ),
       .f = function(a, b) {
         a %>% 
           inner_join(y = b %>% select(time_rnd, wday.lbl.Friday:minute.55),
                      by = c("time_rnd" = "time_rnd")
                      )
         }
       )

message("combo")
combo$`10` %>% glimpse()


rm(small_test, lmtd_vars_lag3)



train_data <-
  combo %>% 
  map(~filter(.x, data_type == "01_train"))

message("train_data")
train_data %>% 
  map(~dim(.x))

```


### Setup for Models Not in `caret` 
  
  This is based on the example shown [here](https://topepo.github.io/rsample/articles/Applications/Time_Series.html), and is needed for the time-series-ish models, which are not currently part of the `caret` modeling process.
    
  Create the rolling-origin resamples to be used for measuring forecast accuracy. Rolling samples will use 0.5 years of data do predict the next 0.25 years of data.  
```{r}

roll_rs <-
  train_data %>% 
  map(~ rolling_origin(.x,
                       initial = period_train,
                       assess = period_test,
                       cumulative = FALSE,
                       skip = skip_span
                       )
      )

message("train_data")
# DV_train_data %>% 
train_data %>% 
  map(~ nrow(.x)
      )
  
message("roll_rs")
roll_rs %>% 
  map(~ dim(.x)
      )

names(roll_rs$`10`)

roll_rs %>% 
  map(~ length(.x$splits)
      )

```
  
    
  For plotting, let’s index each split by the first day of the assessment set.
```{r}

get_date <-
  function(x)
    min(assessment(x)$time_rnd
        )

roll_rs <-
  pmap(.l = list(a = roll_rs),
       .f = function(a) {
         data = a
         
         splits = data$splits %>% 
           map(get_date)
         
         data$start_date = do.call("c", splits)
         
         return(data)
         }
       )

names(roll_rs$`10`)
length(roll_rs$`10`$start_date)
head(roll_rs$`10`$start_date, 20)


rm(get_date)

```


### Model Setup (`forecast::auto.arima` and `prophet::prophet`)
      
  Here, we use `forecast::auto.arima` to produce create an arima model. We also try `prophet::prophet` to create a model based on trend, seasonality, and holidays, and that is a bit like a general additive model. More info can be found [here](https://facebook.github.io/prophet/) and [here](https://peerj.com/preprints/3190/).  
    
  First, I create the function for the basic ARIMA model.
```{r}

fit_model_arima <-
  function(x, ...) {
    data = x %>% 
      analysis() %>% 
      # Since the first day changes over resamples, adjust it based on the first date value in the data frame
      tk_ts(select = bikes_avail_now,
            start = .$time_rnd[[1]] %>% lubridate::year(),
            freq = (12 * 24), # 12 readings per hour for 24 hours
            silent = TRUE
            )
    
    fit = auto.arima(data, ...)
    
    return(fit)
  }

```
  
    
  Next I create the function to run an ARIMA model with external regressors (external regressors include fourier transformations, and other regressors identified by using the random forest and xgboost models used above).
```{r}

fit_model_arima_xreg <-
  function(x, ...) {
    data = 
      x %>% 
      analysis()
    
    # hourly frequency
    ts_hr =
      data %>% 
      tk_ts(select = bikes_avail_now,
            start = .$time_rnd[[1]] %>% lubridate::year(),
            freq = 12, # 12 readings per hour
            silent = TRUE
            )
    
    # daily frequency
    ts_7 =
      data %>% 
      tk_ts(select = bikes_avail_now,
            start = .$time_rnd[[1]] %>% lubridate::year(),
            freq = (12 * 24), # 12 readings per hour for 24 hours
            silent = TRUE
            )
    
    # use a fourier transformation to capture daily seasonality and choose K programatically
    bestfit = list(aicc = Inf)
    for(K in seq(7)
        ) {
      n = nrow(assessment(x)
               )
      
      ts_7_fourier = fourier(ts_7,
                               K = K
                               )
      
      ts_7_fourier_future = fourier(ts_7,
                                      K = K,
                                      h = n
                                      )
      
      fit = auto.arima(ts_hr,
                       xreg = cbind(ts_7_fourier,
                                    # additional variables identified as "important" with Random Forest and XGBTree models
                                    data$wday.lbl,
                                    data$hour,
                                    data$minute,
                                    data$bikes_avail_lag05,
                                    data$bikes_avail_lag10,
                                    data$bikes_avail_lag15
                                    )
                       )
      
      if(fit[["aicc"]] < bestfit[["aicc"]]) {
        bestfit = fit
        bestK = K
        bestts_7_fourier_future = ts_7_fourier_future
      }
      
      return(list(best_fit = bestfit,
                  best_k = bestK,
                  best_ts_7_fourier_future = bestts_7_fourier_future
                  )
             )
    }
    
    return()
  }

```
  
    
  Now I create the function to run the basic prophet model (based just on `el_rides`.
```{r}

# prophet
fit_model_prophet <-
  function(x, ...) {
    
    # Create train, validation, & test sets
        dat_trn_vld = analysis(x)
        
        tot_rows_trn_vld = dat_trn_vld %>% nrow()
    
        splt_rows = round(tot_rows_trn_vld * 0.70)
    
        df_trn = dat_trn_vld[1:splt_rows, , drop = FALSE]
        
        df_val = dat_trn_vld[(splt_rows + 1):tot_rows_trn_vld, , drop = FALSE]
        
        df_tst = assessment(x)
        
        full_data = bind_rows(df_trn %>% add_column(key = "training"),
                              df_val %>% add_column(key = "valid"),
                              df_tst %>% add_column(key = "test")
                              )
        
        pred_rows = nrow(df_val) + nrow(df_tst)
        
        
    # Create the prophet model
        
        prophet =
          df_trn %>% 
          select(time_rnd,
                 bikes_avail_now
                 ) %>% 
          dplyr::rename(ds = time_rnd,
                        y = bikes_avail_now
                        ) %>% 
          prophet(...)
    
        
    # return
        list_return = list(prophet = prophet,
                           pred_rows = pred_rows,
                           full_data = full_data
                           )
        return(list_return)
    
  }

```

  Now I create the function to run the basic prophet model, with the addition of external regressors.
```{r}

# prophet
fit_model_prophet_add_regrs <-
  function(x, ...) {
    dat =
      x %>% 
      analysis() %>% 
      select(time_rnd,
             bikes_avail_now,
             wday.lbl.Monday
             ) %>% 
      dplyr::rename(ds = time_rnd,
                    y = bikes_avail_now
                    )
    
    m = prophet(...)
    
    m = add_regressor(m, "wday.lbl.Monday")
    
    m = fit.prophet(m, dat)
  }

```


### Run Models

#### Arima-Based Model  
  **Need to review this:**  [http://www.business-science.io/code-tools/2018/04/08/introducing-anomalize.html](http://www.business-science.io/code-tools/2018/04/08/introducing-anomalize.html) for anomaly detection.
    
  Can also use any other modeling methods (e.g., prophet), and then use the outlier detection method twitter used (Generalized ESD) on the remainder. **See:**  [https://www.rdocumentation.org/packages/EnvStats/versions/2.3.0/topics/rosnerTest](https://www.rdocumentation.org/packages/EnvStats/versions/2.3.0/topics/rosnerTest) for the algorithm used outside of `AnomalyDetection::AnomalyDetectionTs`.  
    
  Here I run the basic `forecast::auto.arima` model.
```{r}

# user   system  elapsed 
# 1001.467   21.414 1036.904
# ~ 17 min
message("arima")
# start <- proc.time()
# models <-
#   pmap(.l = list(a = roll_rs),
#        .f = function(a) {
# 
#          splits_a = a$splits %>%
#            map(fit_model_arima)
# 
#          a$arima = splits_a
# 
#          return(a)
#          }
#        )

arima_1_split <-
  roll_rs$`10`$splits[[1]] %>% fit_model_arima()
arima_1_split

time.arima <- proc.time() - start
time.arima

```
  
    
  Here I run the `forecast::auto.arima` model with external regressors.
```{r}
 
# user  system elapsed 
# 626.467  26.937 673.471
# ~ 11 min
message("arima_xreg")
start <- proc.time()
# models <-
#   pmap(.l = list(a = models),
#        .f = function(a) {
# 
#          splits_a_xreg = a$splits %>%
#            map(fit_model_arima_xreg)
# 
#          a$arima_xreg = splits_a_xreg
# 
#          return(a)
#          }
#        )

arima_xreg_1_split <-
  roll_rs$`10`$splits[[1]] %>% fit_model_arima_xreg()
arima_xreg_1_split

time.arima_xreg <- proc.time() - start
time.arima_xreg
```
  
    
  Here I run the basic `prophet::prophet` model
```{r}

# user   system  elapsed 
# 3223.964    5.736 3228.870
# ~ 54 minutes for two station_id values

message("prophet")
start <- proc.time()
mod_prophet <-
  pmap(.l = list(a = roll_rs),
       .f = function(a) {

         splits_p = a$splits %>%
           map(fit_model_prophet)

         a$prophet_mod = splits_p

         return(a)
         }
       )

time.prophet <- proc.time() - start
time.prophet


saveRDS(mod_prophet,
        paste0(wd,
               "/Models/",
               "mod_prophet.Rds"
               )
        )

# mod_prophet <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "mod_prophet.Rds"
#                  )
#           )

```
  
    
  Here I run the `prophet::prophet` model that includes an additional regressor for "Monday" as a test.
```{r}

# user  system elapsed 
#  35.327   3.401  47.577
message("prophet_hol")
start <- proc.time()
# models <-
#   pmap(.l = list(a = models),
#        .f = function(a) {
# 
#          splits_p_hol = a$splits %>%
#            map(fit_model_prophet_add_regrs)
# 
#          a$prophet_hol <- splits_p_hol
# 
#          return(a)
#          }
#        )

prophet_add_mon_regressor <-
  roll_rs$`10`$splits[[1]] %>% fit_model_prophet_add_regrs()
prophet_add_mon_regressor

time.prophet_hol <- proc.time() - start
time.prophet_hol

```



```{r}

run_times <- list(arima = as.list(time.arima),
                  arima_xreg = as.list(time.arima_xreg),
                  prophet = as.list(time.prophet),
                  prophet_hol = as.list(time.prophet_hol)
                  )
run_times

# saving is done to avoid having to run the models again
saveRDS(models,
        paste0(wd,
               "/Models/",
               "models.Rds"
               )
        )
# models <-
#   readRDS(paste0(wd,
#                  "/Data_Processed/",
#                  "models.Rds"
#                  )
#           )

```
  
    
  Create the `Prophet` forecasts and plots - these will be used for measuring model accuracy (below). Here I create the future prophet dataset.
```{r}

mod_prophet <-
  pmap(.l = list(a = mod_prophet),
       .f = function(a) {
         
         splits_p_future =
           pmap(.l = list(b = a$prophet_mod),
                .f = function(b) {
                  make_future_dataframe(m = b$prophet,
                                        periods = b$pred_rows,
                                        freq = (60 * 5) # 5 min intervals
                                        )
                  }
                )
         
         a$prophet.future <- splits_p_future
         
         return(a)
         }
       )

length(mod_prophet$`10`$prophet.future)

```  
  
    
  And now I can create the prohpet forecasts. First, the basic prophet forecast.
```{r}

# user    system   elapsed 
# 15395.248 22280.912  7883.235
# ~ 131 min
start <- proc.time()
mod_prophet <-
  pmap(.l = list(a = mod_prophet),
       .f = function(a) {
         splits_p_m =
           a$prophet_mod
         
         splits_p_future =
           a$prophet.future
         
         splits_p_forecast =
           pmap(.l = list(b = splits_p_m,
                          c = splits_p_future
                          ),
                .f = function(b, c) {
                  predict(b$prophet, c)
                  }
                )
         
         a$prophet.forecast = splits_p_forecast
         
         return(a)
         }
       )

time.prophet.forecast <- proc.time() - start
time.prophet.forecast


saveRDS(mod_prophet,
        paste0(wd,
               "/Models/",
               "mod_prophet.Rds"
               )
        )

# mod_prophet <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "mod_prophet.Rds"
#                  )
#           )

```
  
    
  And now the prophet forecast including holidays.
```{r}

# user  system elapsed 
# 792.629  78.698 881.271
# ~ 15 min
start <- proc.time()
prophet.forecast <-
  pmap(.l = list(a = prophet.forecast),
       .f = function(a) {
         splits_p_hol_m =
           a$prophet_hol
         
         splits_p_hol_future =
           a$prophet_hol.future
         
         splits_p_hol_forecast =
           pmap(.l = list(b = splits_p_hol_m,
                          c = splits_p_hol_future
                          ),
                .f = function(b, c) {
                  predict(b, c) %>% 
                    # if_else is needed to prevent any negative predictions
                    mutate(yhat_zero_floor = if_else(yhat < 0,
                                                     0,
                                                     yhat
                                                     )
                           )
                  }
                )
         
         a$prophet_hol.forecast = splits_p_hol_forecast
         
         return(a)
         }
       )
time.prophet_hol.forecast <- proc.time() - start
time.prophet_hol.forecast

```
  
    
  Now I simply update the `run_times` dataset with the relevant prophet info, and the relvant info from the random forest and xgboost models.
```{r}

run_times[5:6] <-
  list(prophet.forecast = as.list(time.prophet.forecast),
       prophet_hol.forecast = as.list(time.prophet_hol.forecast)
       )
names(run_times)[5:6] <- c("prophet.forecast", "prophet_hol.forecast")
run_times[7:8] <-
  list(rf_corr_no = as.list(time.Rf.corr_no),
       xgbtree_corr_yes = as.list(time.Xgbtree.corr_yes)
       )
  
names(run_times)[7:8] <- c("rf_corr_no", "xgbtree_corr_yes")
str(run_times)

# saving is done to avoid having to run the forecasts again
saveRDS(run_times,
        paste0(wd,
               "/Models/",
               "run_times.Rds"
               )
        )

# run_times <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "run_times.Rds"
#                  )
#           )

rm(time.Rf.corr_no, time.Xgbtree.corr_yes)

```
  
    
  And now I can create the prophet plots of components.
```{r}

prophet_plot_components(mod_prophet$`10`$prophet_mod[[1]]$prophet,
                        mod_prophet$`10`$prophet.forecast[[1]]
                        )

```
  
    
  Calculating the error metrics on the validation data. First, we create the by-row error.
```{r}

mod_prophet <-
  pmap(.l = list(a = mod_prophet),
       .f = function(a) {
         p_mod = a$prophet_mod
         fcast = a$prophet.forecast
         
         error_tbl =
           pmap(.l = list(b = p_mod,
                          c = fcast
                          ),
                .f = function(b, c) {
                  tbl_valid = b$full_data %>% 
                    filter(key == "valid") %>% 
                    select(station_id, time_rnd, bikes_avail_now)
                  
                  preds = c %>% select(ds, yhat)
                  
                  error_tbl = tbl_valid %>% 
                    left_join(y = preds,
                              by = c("time_rnd" = "ds")
                              ) %>% 
                    dplyr::rename(actual = bikes_avail_now,
                                  pred = yhat
                                  ) %>% 
                    mutate(pred_rnd = round(pred),
                           error = actual - pred_rnd,
                           error_pct = error / actual
                           )
                  
                  return(error_tbl)
                  }
                )
         
         a$error_tbl = error_tbl
         
         return(a)
         }
       )


saveRDS(mod_prophet,
        paste0(wd,
               "/Models/",
               "mod_prophet.Rds"
               )
        )

# mod_prophet <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "mod_prophet.Rds"
#                  )
#           )

```
  
    
  Next, we create the summary stats of the error metrics.
```{r}

error_sum_stats <-
  pmap(.l = list(a = mod_prophet),
       .f = function(a) {
         error_tbl = a$error_tbl
         
         sum_stats =
           pmap(.l = list(b = error_tbl),
                .f = function(b) {
                  b %>% 
                    summarise(mae = mean(abs(error), na.rm = TRUE),
                              mse = mean(error^2, na.rm = TRUE),
                              rmse = sqrt(mse),
                              mape = mean(abs(error_pct), na.rm = TRUE)
                              )
                  }
                )
         }
       )


error_sum_stats_mean <-
  pmap(.l = list(a = error_sum_stats,
                 b = names(error_sum_stats)
                 ),
       .f = function(a, b) {
         df = bind_rows(a)
         
         means = df %>% 
           summarise(station_id = b,
                     mae_mean = mean(mae, na.rm = TRUE),
                     mse_mean = mean(mse, na.rm = TRUE),
                     rmse_mean = mean(rmse, na.rm = TRUE),
                     mape_mean = mean(mape, na.rm = TRUE)
                     )
         }
       )

error_sum_stats_mean %>% 
  bind_rows()


saveRDS(error_sum_stats_mean,
        paste0(wd,
               "/Models/",
               "error_sum_stats_mean.Rds"
               )
        )

# error_sum_stats_mean <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "error_sum_stats_mean.Rds"
#                  )
#           )

```
  
    
  Looking at an example "actual" vs. "pred" plot.
```{r}

mod_prophet$`10`$error_tbl[[10]]$time_rnd %>% min()
mod_prophet$`10`$error_tbl[[10]]$time_rnd %>% max()


mod_prophet$`10`$error_tbl[[10]] %>% 
  gather(key = "key",
         value = "value",
         actual,
         pred_rnd
         ) %>% 
  ggplot(aes(x = time_rnd, y = value, color = key)) +
        geom_line() +
        coord_cartesian(
          xlim = c(as_datetime("2014-09-06 00:00:00",
                                             tz = "America/Los_Angeles"
                                             ),
                                 as_datetime("2014-09-08 00:00:00",
                                             tz = "America/Los_Angeles"
                                             )
                                 ),
                        ylim = c(0, 8)
                        ) +
        scale_y_continuous(breaks = seq(0, 8, 2)) +
        theme_minimal() +
        theme(legend.position = "bottom")


```


