---
title: "SF Bikeshare Status - Modeling With H2O"
output: html_notebook
---


  
**NOTE:** before running this chunk, the AWS Instance Type was switched from t2.micro to t3.2xlarge.


## Setup    
  Load the relevant libraries.
```{r}

# {r, message=FALSE, warning=FALSE}
# rm(list = ls())
# .rs.restartR()

library("tidyverse")
# install.packages("data.table")
library("data.table")
library("rsample")
# library("lubridate")
# install.packages("timetk")
# library("timetk")
# library("DBI")
# install.packages("RSQLite")
# library("RSQLite")
# install.packages("rJava")
# library("rJava")
# install.packages("h2o")
library("h2o")
# install.packages("tibble")
library("furrr")
library("glue")
# install.packages("reticulate")
library("reticulate")

```
  
    
  Session Info.
```{r}

sessionInfo()

```


  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/Bikeshare_Status_SF/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
  
    
## Setup for running processes (i.e. models) in parallel on EC2.  
  
 Basic info needed for the setup.
```{r}
# # A t3.2xlarge AWS instance
# # Created from http://www.louisaslett.com/RStudio_AMI/
# public_ip <- "34.229.223.90"
# 
# # This is where my pem file lives (password file to connect).
# ssh_private_key_file <- "/home/rstudio/Dropbox/_AWS/Dot_Pem/LouisAslett_AMI_Key_Pair.pem"
```
  
    
  Make the cluster.
```{r}
# cl <- makeClusterPSOCK(
#   
#   # Public IP number of EC2 instance
#   workers = public_ip,
#   
#   # User name (always 'ubuntu')
#   user = "ubuntu",
#   
#   # Use private SSH key registered with AWS
#   rshopts = c(
#     "-o", "StrictHostKeyChecking=no",
#     "-o", "IdentitiesOnly=yes",
#     "-i", ssh_private_key_file
#   ),
#   
#   # Set up .libPaths() for the 'ubuntu' user and
#   # install furrr
#   rscript_args = c(
#     "-e", shQuote("local({p <- Sys.getenv('R_LIBS_USER'); dir.create(p, recursive = TRUE, showWarnings = FALSE); .libPaths(p)})"),
#     "-e", shQuote("install.packages(c('furrr', 'h2o'))")
#   ),
#   
#   # Switch this to TRUE to see the code that is run on the workers without
#   # making the connection
#   dryrun = FALSE
# )
# 
# cl
```


```{r}
# boto <- import("boto3")
# 
# # Import EC2 client
# ec2 <- boto$client('ec2')
# ec2
```



```{r}
# h2o.makeEC2ClusterPsock <-
#   function(public_ip, ssh_private_key_file, max_mem_size = "10g", dryrun = F, h2o = T){
# 
#   if(h2o){
#     r_command <- c(
#       "-e", shQuote(".libPaths('/home/rstudio/R/x86_64-pc-linux-gnu-library/3.5')"),
#       "-e", shQuote(glue("if (!require('furrr')) install.packages('furrr');library(h2o);system('pkill java');h2o.init(nthreads = -1, max_mem_size = '{max_mem_size}');h2o.removeAll();gc()"))
#     )
#   } else {
#     r_command <- c(
#       "-e", shQuote(".libPaths('/home/rstudio/R/x86_64-pc-linux-gnu-library/3.5')"),
#       "-e", shQuote("if (!require('furrr')) install.packages('furrr')")
#     )
#    }
# 
#   makeClusterPSOCK(
# 
#     # Public IP number of EC2 instance
#     workers = public_ip,
# 
#     # User name (always 'ubuntu')
#     user = "ubuntu",
# 
#     # Use private SSH key registered with AWS
#     # rshcmd = c("plink", "-ssh", "-i", ssh_private_key_file),
#     rshopts = c(
#     "-o", "StrictHostKeyChecking=no",
#     "-o", "IdentitiesOnly=yes",
#     "-i", ssh_private_key_file
#   ),
# 
#     # Check available memory
#     # Set up .libPaths() for the 'ubuntu' user and
#     # install furrr
#     rscript_args = r_command,
# 
#     # Switch this to TRUE to see the code that is run on the workers without
#     # making the connection
#     dryrun = dryrun, 
#     verbose = T
#   )
# }
```


```{r}
# time_start <- proc.time()
# 
# plan(cluster, workers = cl)
# 
# cars2_mod_future <- mtcars %>%
#   split(.$gear) %>%
#   future_map(~lm(mpg ~ cyl + hp + wt, data = .))
#   # map(~lm(mpg ~ cyl + hp + wt, data = .))
# 
# time_end <- proc.time()
# 
# time_end - time_start
# 
# rm(time_start, time_end)
```


## Modeling  
  
  Get the base dataset.
```{r}
# id_split <-
#   read_rds(path = paste0(wd,
#                          "/Data/Interim/",
#                          "id_split.rds"
#                          )
#            )
# pmap(.l = list(a = id_split,
#                b = names(id_split)
#                ),
#      .f = function(a, b) {
#        message(b)
#        
#        str(a)
#        }
#      )
# View(id_split$`10` %>% head(1000))

lmtd_vars_lag3 <- lmtd_vars_lag3 %>% 
  map(~select(.x,
              station_id:minute.55
              )
      )

names(lmtd_vars_lag3)
lmtd_vars_lag3$`10` %>% class()
lmtd_vars_lag3$`10` %>% glimpse()

```
  
    
  Create splits for modeling.
```{r}

# period_train <- 6 * 24 * 365 * 1.5 # 1.5 years of training
# period_test  <- 6 * 24 * 365 * 0.5 # 0.5 year of training
# skip_span     <- 6 * 24 * 75 # produces 10 splits
# # period_train <- (6 * 24 * 365 * 1) # 1.0 years of training data
# # period_test <- (6 * 24 * 365 * 0.5) # 0.5 years of testing data
# # skip_span <- (6 * 24 * 50) # gives 10 evenly distributed  splits
# 
# rolling_origin_resamples <- 
#   lmtd_vars_lag3 %>% 
#   map(~rolling_origin(.x,
#                       initial    = period_train,
#                       assess     = period_test,
#                       cumulative = FALSE,
#                       skip       = skip_span
#                       )
#       )
# 
# rolling_origin_resamples %>% map(~nrow(.x))
# 
# rolling_origin_resamples <- rolling_origin_resamples[1:2]
# 
# names(rolling_origin_resamples)
# rolling_origin_resamples$`10`$splits[[10]] %>% analysis() %>% class()
# rolling_origin_resamples$`10`$splits[[10]] %>% analysis() %>% glimpse()

```
  

# Run a h2o::h2o.automl for each split for each station_id.

```{r}

train_data <-
  lmtd_vars_lag3 %>% 
  map(~filter(.x, data_type == "01_train"))


num_rows <- train_data %>% map(~nrow(.x))
splt_pt_train <- num_rows %>% map(~round((.x * 0.70), digits = 0))
splt_pt_valid <- num_rows %>% map(~round((.x * 0.85), digits = 0))


tbl_train <-
  pmap(.l = list(a = train_data,
                 b = splt_pt_train
                 ),
       .f = function(a, b) {
         dat = a[1:b, ] %>% 
           select(-station_id,
                  -data_type,
                  -time_rnd,
                  -row_num
                  )
         }
       )

tbl_valid <-
  pmap(.l = list(a = train_data,
                 b = splt_pt_train,
                 c = splt_pt_valid
                 ),
       .f = function(a, b, c) {
         dat = a[(b + 1):c, ] %>% 
           select(-station_id,
                  -data_type,
                  -time_rnd,
                  -row_num
                  )
         }
       )

tbl_test <-
  pmap(.l = list(a = train_data,
                 b = splt_pt_valid,
                 c = num_rows
                 ),
       .f = function(a, b, c) {
         dat = a[(b + 1):c, ] %>% 
           select(-station_id,
                  -data_type,
                  -time_rnd,
                  -row_num
                  )
         }
       )


rm(num_rows, splt_pt_train, splt_pt_valid)


# valid_data is what will be used to compare all model types
valid_data <-
  lmtd_vars_lag3 %>% 
  map(~filter(.x, data_type == "02_valid") %>% 
        select(-station_id,
               -data_type,
               -time_rnd,
               -row_num
               )
      )

```
  
    
  Start h2o.
```{r}

h2o.init()
h2o.no_progress() # Turn off progress bars

```
  
    
  Convert to the datasets to h2o objects.
```{r}

train_h2o <- tbl_train %>% map(~as.h2o(.x))
valid_h2o <- tbl_valid %>% map(~as.h2o(.x))
test_h2o  <- tbl_test %>% map(~as.h2o(.x))

valid_data_h2o  <- valid_data %>% map(~as.h2o(.x))


# rm(tbl_train, tbl_valid, tbl_test)

```
  
    
  Set the relevant variable names.
```{r}

y <- "bikes_avail_now"

x <- train_h2o %>% 
  map(~setdiff(names(.x),
               y
               )
      )

```
  
    
  Run `h2o.automl`.
```{r}

# h2o.shutdown(prompt = TRUE)
# plan(multiprocess)
# cl <- h2o.makeEC2ClusterPsock(public_ip = public_ip,
#                               max_mem_size = "5g",
#                               ssh_private_key_file = ssh_private_key_file,
#                               dryrun = F,
#                               h2o = T
#                               )
# plan(cluster, workers = cl)


# user   system  elapsed 
#   82.212   26.563 7291.869 
# ~ 2 hours
start <- proc.time()

automl_models_h2o <-
  pmap(
  # future_pmap(
    .l = list(
      a = x,
      b = train_h2o,
      c = valid_h2o,
      d = test_h2o
      ),
    .f = function(a, b, c, d) {
      h2o.automl(x = a,
                    y = y,
                    training_frame = b,
                    validation_frame = c,
                    leaderboard_frame = d,
                    nfolds = 10,
                    max_runtime_secs = 3600, # 1 hour
                    # max_runtime_secs = 60, # 1 hour
                    max_models = 10,
                    stopping_metric = "deviance",
                    seed = 123456789
                    )
         }
       )

h2o.time <- proc.time() - start
h2o.time


rm(start)

```
  
    
  Save the resulting .rds files.
```{r}

saveRDS(h2o.time,
        paste0(wd,
               "/Models/",
               "h2o.time.rds"
               )
        )

# h2o.time <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "h2o.time.rds"
#                          )
#            )

saveRDS(automl_models_h2o,
        paste0(wd,
               "/Models/",
               "automl_models_h2o.rds"
               )
        )

# automl_models_h2o <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "automl_models_h2o.rds"
#                          )
#            )

```
  
    
  Extract leader models, and save each model.
```{r}

# extract the leader models
automl_leader <-
  pmap(.l = list(a = automl_models_h2o),
       .f = function(a) {
         a@leader
         }
       )

# str(automl_models_h2o)


# save each individual model
model_path_h2o <-
  automl_leader %>% 
  map(~h2o.saveModel(object = .x,
                     path = paste0(wd,
                                   "/Models/"
                                   ),
                     force = TRUE
                     )
      )

saveRDS(model_path_h2o,
        paste0(wd,
               "/Models/",
               "model_path_h2o.rds"
               )
        )

# model_path_h2o <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "model_path_h2o.rds"
#                          )
#            )
# 
# print(model_path_h2o)

# load the model
# automl_leader <-
#   model_path_h2o %>%
#   map(~h2o.loadModel(.x)
#       )

```
  
    
  Investigate variable importance of the leader models.
```{r}

leader_models_varimp <-
  pmap(.l = list(a = automl_models_h2o),
       .f = function(a) {
         h2o.varimp(object = a@leader)
         }
       )

leader_models_varimp_plot <-
  pmap(.l = list(a = automl_models_h2o),
       .f = function(a) {
         h2o.varimp_plot(model = a@leader, num_of_features = 10)
         }
       )

leader_models_varimp %>% 
  map(~as.data.frame(.x))

leader_models_varimp_plot

```
  
    
  Create predictions from the models.
```{r}

pred_h2o <-
  pmap(.l = list(a = automl_leader,
                 b = test_h2o
                 ),
       .f = function(a, b) {
         h2o.predict(a, newdata = b)
         }
       )

# save as a .rds file
saveRDS(pred_h2o,
        paste0(wd,
               "/Models/",
               "pred_h2o.rds"
               )
        )

# pred_h2o <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "pred_h2o.rds"
#                          )
#            )

```
  
    
  Look at performance stats using the test data.
```{r}

pmap(.l = list(a = automl_leader,
               b = test_h2o
               ),
     .f = function(a, b) {
       h2o.performance(a, newdata = b)
       }
     )

```
  
    
  Investigate test error.
```{r}

error_tbl_h2o <-
  pmap(.l = list(a = tbl_test,
                 b = pred_h2o
                 ),
       .f = function(a, b) {
         a %>% 
           add_column(pred = b %>% as_tibble() %>% pull(predict)) %>% 
           rename(actual = bikes_avail_now) %>% 
           mutate(error = actual - pred,
                  error_pct = error / actual
                  ) 
         }
       )


# save as a .rds file
saveRDS(error_tbl_h2o,
        paste0(wd,
               "/Models/",
               "error_tbl_h2o.rds"
               )
        )

# error_tbl_h2o <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "error_tbl_h2o.rds"
#                          )
#            )

```


```{r}

pmap(.l = list(a = error_tbl_h2o,
               b = names(error_tbl_h2o)
               ),
     .f = function(a, b) {
       a %>% 
         summarise(me   = mean(error, na.rm = TRUE),
                   rmse = mean(error^2, na.rm = TRUE)^0.5,
                   mae  = mean(abs(error), na.rm = TRUE),
                   mape = mean(abs(error_pct), na.rm = TRUE),
                   mpe  = mean(error_pct, na.rm = TRUE)
                   ) %>% 
          # glimpse() %>% 
         mutate(station_id = b)
       }
     ) %>% 
  bind_rows()

```
  
    
  Create predictions from the models on the validation data.
```{r}

valid_data_pred_h2o <-
  pmap(.l = list(a = automl_leader,
                 b = valid_data_h2o
                 ),
       .f = function(a, b) {
         h2o.predict(a, newdata = b)
         }
       )

# save as a .rds file
saveRDS(valid_data_pred_h2o,
        paste0(wd,
               "/Models/",
               "valid_data_pred_h2o.rds"
               )
        )

# valid_data_pred_h2o <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "valid_data_pred_h2o.rds"
#                          )
#            )

```
  
    
  Look at performance stats using the test data.
```{r}

pmap(.l = list(a = automl_leader,
               b = valid_data_h2o
               ),
     .f = function(a, b) {
       h2o.performance(a, newdata = b)
       }
     )

```
  
    
  Investigate test error.
```{r}

valid_data_error_tbl_h2o <-
  pmap(.l = list(a = valid_data,
                 b = valid_data_pred_h2o
                 ),
       .f = function(a, b) {
         a %>% 
           add_column(pred = b %>% as_tibble() %>% pull(predict)) %>% 
           rename(actual = bikes_avail_now) %>% 
           mutate(error = actual - pred,
                  error_pct = error / actual
                  ) 
         }
       )


# save as a .rds file
saveRDS(valid_data_error_tbl_h2o,
        paste0(wd,
               "/Models/",
               "valid_data_error_tbl_h2o.rds"
               )
        )

# valid_data_error_tbl_h2o <-
#   read_rds(path = paste0(wd,
#                          "/Models/",
#                          "valid_data_error_tbl_h2o.rds"
#                          )
#            )

```


```{r}

pmap(.l = list(a = valid_data_error_tbl_h2o,
               b = names(valid_data_error_tbl_h2o)
               ),
     .f = function(a, b) {
       a %>% 
         summarise(me   = mean(error, na.rm = TRUE),
                   rmse = mean(error^2, na.rm = TRUE)^0.5,
                   mae  = mean(abs(error), na.rm = TRUE),
                   mape = mean(abs(error_pct), na.rm = TRUE),
                   mpe  = mean(error_pct, na.rm = TRUE)
                   ) %>% 
          # glimpse() %>% 
         mutate(station_id = b)
       }
     ) %>% 
  bind_rows()

```
    
    
  Remove no-longer-needed files.
```{r}

rm(automl_leader, automl_models_h2o, lmtd_vars_lag3, leader_models_varimp, leader_models_varimp_plot, model_path_h2o, pred_h2o, test_h2o, train_h2o, valid_h2o, x, h2o.time, y
   )

```
  
    
  Shutting down h2o.
```{r}

h2o.shutdown()

```
  
  
