---
title: "Step 04: Models - Lag 5min Model"
output:
  html_document:
    df_print: paged
---
  
## Setup    
  Load the relevant libraries.
```{r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()

# data manipulation
library("plyr")
library("tidyverse")
library("magrittr")

# other
library("rsample")            # rolling samples for validation stats

```
  
    
  Session Info.
```{r}

sessionInfo()

```
  
    
  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/Bikeshare_Status_SF/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```


## Modeling

**NOTE: `period_train`, `period_test`, and `skip_span`, are the outputs produced in Step 02 and Step 03**
```{r}

period_train <- # (12 * 24 * 365 * 0.5) # 0.5 years of training data
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "period_train.Rds"
                 )
          )

period_test <- # (12 * 24 * 365 * 0.25) # 0.25 years of testing data
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "period_test.Rds"
                 )
          )

skip_span <- # (12 * 24 * 25) # gives 10 evenly distributed  splits
  readRDS(paste0(wd,
                 "/Data/Interim/",
                 "skip_span.Rds"
                 )
          )

```  
  
    
  Creating the datasets used in modeling.
```{r}

message("small_test")
names(small_test)
small_test$`10` %>% class()
small_test$`10` %>% glimpse()



train_data <-
  small_test %>% 
  map(~filter(.x, data_type == "01_train"))

message("train_data")
train_data %>% 
  map(~dim(.x))


rm(lmtd_vars_lag3)

```


### Setup for Models Not in `caret` 
  
  This is based on the example shown [here](https://topepo.github.io/rsample/articles/Applications/Time_Series.html), and is needed for the time-series-ish models, which are not currently part of the `caret` modeling process.
    
  Create the rolling-origin resamples to be used for measuring forecast accuracy. Rolling samples will use 0.5 years of data do predict the next 0.25 years of data.  
```{r}

roll_rs <-
  train_data %>% 
  map(~ rolling_origin(.x,
                       initial = period_train,
                       assess = period_test,
                       cumulative = FALSE,
                       skip = skip_span
                       )
      )

message("train_data")
# DV_train_data %>% 
train_data %>% 
  map(~ nrow(.x)
      )
  
message("roll_rs")
roll_rs %>% 
  map(~ dim(.x)
      )

names(roll_rs$`10`)

roll_rs %>% 
  map(~ length(.x$splits)
      )

```
  
    
  For plotting, let’s index each split by the first day of the assessment set.
```{r}

get_date <-
  function(x)
    min(assessment(x)$time_rnd
        )

roll_rs <-
  pmap(.l = list(a = roll_rs),
       .f = function(a) {
         data = a
         
         splits = data$splits %>% 
           map(get_date)
         
         data$start_date = do.call("c", splits)
         
         return(data)
         }
       )

names(roll_rs$`10`)
length(roll_rs$`10`$start_date)
head(roll_rs$`10`$start_date, 20)


rm(get_date)

```


### Model Setup
  
  Here I create the function to run the basic lag1 (e.g., lag of 5 min) model to use as a baseline.
```{r}

fit_model_lag_05min <-
  function(x, ...) {
    
    # Run just for the test data (for comparison with other models)
        df_tst = assessment(x)
        
        lag_5min =
          df_tst %>% 
          select(station_id,
                 time_rnd,
                 bikes_avail_now,
                 bikes_avail_lag05
                 ) %>% 
          dplyr::rename(actual = bikes_avail_now,
                        lag_05 = bikes_avail_lag05
                        )
    
        return(lag_5min)
  }

```


### Run Models

#### Arima-Based Model  
  
    
  Here I run the basic lag model.
```{r}

# user  system elapsed 
#   0.552   0.032   0.589 
start <- proc.time()
mod_lag_05 <-
  pmap(.l = list(a = roll_rs),
       .f = function(a) {

         splits_p = a$splits %>%
           map(fit_model_lag_05min)

         a$lag_5min = splits_p

         return(a)
         }
       )

time.lag_05 <- proc.time() - start
time.lag_05


saveRDS(mod_lag_05,
        paste0(wd,
               "/Models/",
               "mod_lag_05.Rds"
               )
        )

# mod_lag_05 <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "mod_lag_05.Rds"
#                  )
#           )


saveRDS(time.lag_05,
        paste0(wd,
               "/Models/",
               "time.lag_05.Rds"
               )
        )

# time.lag_05 <-
#   readRDS(paste0(wd,
#                  "/Models/",
#                  "time.lag_05.Rds"
#                  )
#           )


rm(roll_rs)

```
  
    
  Remove no-longer-needed files.
```{r}

rm(small_test, train_data, period_test, period_train, skip_span, start, fit_model_lag_05min)

```


