---
title: "SF Bikeshare Status - Modeling Compare"
output: html_notebook
---


**NOTE:** before running this chunk, the AWS Instance Type was switched from t2.micro to t3.2xlarge.

**NOTE:** `model_setup.Rmd` should be run first to create the needed datasets for the model comparison.



## Setup    
  Load the relevant libraries.
```{r}
# {r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()

# data manipulation
library("tidyverse")
library("data.table")
library("lubridate")


# plotting graphs as .gif and .mpeg files
library("gganimate")
library("gifski")
library("transformr")
library("av")

```
  
    
  Session Info.
```{r}

sessionInfo()

```


  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

# Use below when running from AWS
# opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/Bikeshare_Status_SF/")

# Use below when running from laptop
opts_knit$set(root.dir = "/Users/mdturse/Dropbox/_AWS/Bikeshare_Status_SF/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
  
    
  Remove unneeded files.
```{r}

rm(lmtd_vars_lag3)

```
  
    
  Get the base datasets.
```{r}

error_tbl_h2o <-
  read_rds(path = paste0(wd,
                         "/Models/",
                         "error_tbl_h2o.rds"
                         )
           )


keras_predict_all_stations_all_splits <-
  read_rds(path = paste0(wd,
                         "/Models/",
                         "keras_predict_all_stations_all_splits.rds"
                         )
           )


mod_prophet <-
  readRDS(paste0(wd,
                 "/Models/",
                 "mod_prophet.Rds"
                 )
          )

mod_prophet_error_tbl <-
  pmap(.l = list(a = mod_prophet),
       .f = function(a) {
         error_tbl = a$error_tbl
         return(error_tbl)
         }
       )

rm(mod_prophet)


mod_lag_05 <-
  readRDS(paste0(wd,
                 "/Models/",
                 "mod_lag_05.Rds"
                 )
          )
  
```
  
    
  Create a single dataset with actual values and predicted values (from all the models).
```{r}

all_models_preds <-
  pmap(.l = list(a = keras_predict_all_stations_all_splits,
                 b = mod_prophet_error_tbl,
                 c = error_tbl_h2o,
                 d = mod_lag_05
                 ),
       .f = function(a, b, c, d) {
         h2o_test_preds = c
         
         full_preds =
           pmap(.l = list(e = a$actual_vs_pred,
                          f = b,
                          g = d$lag_5min
                          ),
                .f = function(e, f, g) {
                  keras = e %>% 
                    select(station_id,
                           time_rnd,
                           bikes_avail_now,
                           keras_lstm_uns,
                           keras_lstm_uns_rnd
                           )
                  
                  prophet = f %>% 
                    select(station_id,
                           time_rnd,
                           pred,
                           pred_rnd
                           ) %>% 
                    dplyr::rename(prophet = pred,
                                  prophet_rnd = pred_rnd
                                  )
                  
                  h2o = h2o_test_preds %>%
                    select(station_id,
                           time_rnd,
                           pred
                           ) %>%
                    dplyr::rename(h2o = pred) %>%
                    mutate(h2o_rnd = round(h2o))
                  
                  lag_05 = g
                  
                  full_data =
                    keras %>% 
                    inner_join(y = prophet,
                               by = c("station_id" = "station_id",
                                      "time_rnd" = "time_rnd"
                                      )
                               ) %>% 
                    inner_join(y = h2o,
                               by = c("station_id" = "station_id",
                                      "time_rnd" = "time_rnd"
                                      )
                               ) %>%
                    inner_join(y = lag_05,
                               by = c("station_id" = "station_id",
                                      "time_rnd" = "time_rnd"
                                      ) 
                               ) %>% 
                    select(station_id,
                           time_rnd,
                           bikes_avail_now,
                           keras_lstm_uns_rnd,
                           prophet_rnd,
                           h2o_rnd,
                           lag_05
                           ) %>% 
                    dplyr::rename(actual = bikes_avail_now,
                                  keras = keras_lstm_uns_rnd,
                                  prophet = prophet_rnd,
                                  h2o = h2o_rnd
                                  )
                  
                  return(full_data)
                  }
                )
         
         return(full_preds)
         }
       )


all_models_preds$`10` %>% 
  map(~filter(.x,
              !is.na(keras) &
                !is.na(prophet) &
                !is.na(h2o)
              ) %>% 
        summary()
      )

all_models_preds$`10` %>% 
  map(~summary(.x)
      )

```
  
    
  Basic summary stats of the predictions.
```{r}

all_models_preds_single_df <-
  all_models_preds %>% 
  map(~bind_rows(.x)
      )

all_models_preds_single_df$`61` %>% 
  filter(!is.na(keras) &
           !is.na(prophet) &
           !is.na(h2o) &
           !is.na(lag_05)
         ) %>% 
  summary()

```  
  
    
  Create the basic error statistics (e.g., RMSE, MAE, etc.).
```{r}

error_stats <-
  pmap(.l = list(a = all_models_preds
                 ),
       .f = function(a) {
         data = a %>% 
           bind_rows() %>% 
           # this filter is needed as not all split-model combinations predict the same data
           filter(!is.na(keras) &
                    !is.na(prophet) &
                    !is.na(h2o) &
                    !is.na(lag_05)
                  ) %>%
           gather(key = "model",
                  value = "pred",
                  keras,
                  prophet,
                  h2o,
                  lag_05
                  ) %>% 
           mutate(error = actual - pred,
                  error_pct = error / actual
                  ) %>% 
           group_by(station_id,
                    model
                    ) %>%
           dplyr::summarise(me = mean(error, na.rm = TRUE),
                            rmse = mean(error^2, na.rm = TRUE)^0.5,
                            mae  = mean(abs(error), na.rm = TRUE),
                            mape = mean(abs(error_pct), na.rm = TRUE),
                            mpe  = mean(error_pct, na.rm = TRUE)
                            ) %>% 
           ungroup() %>%
           arrange(station_id,
                   model
                   )
         
         return(data)
         }
       )

# error_stats %>% map(~str(.x))
# error_stats$`10` %>% View()

# error_stats %>% bind_rows %>% View()

```
  
    
  Plot the MAE for each model, for each station_id, then save the graph.
```{r}

mae_bar_graph <-
  error_stats %>% 
  bind_rows() %>% 
  ggplot(aes(x = model,
             y = mae,
             fill = model
             )
         ) +
  geom_col() +
  geom_text(aes(label = format(round(mae,
                                     digits = 1
                                     ),
                               nsmall = 1
                               )
                ),
            hjust = 0.1,
            size = 2
            ) +
  facet_wrap(~station_id
             ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Model Comparison",
       subtitle = "Mean Absolute Error (MAE)",
       x = "Model",
       y = "MAE"
       ) +
  NULL

mae_bar_graph


ggsave(filename = "mae_bar_graph.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```
  
    
  Plot an example of the actual vs. predictions over time for `station_id` = 10.
```{r}

all_models_preds$`10`[[10]] %>%
  filter(!is.na(keras) &
           !is.na(prophet) &
           !is.na(h2o)
         ) %>% 
  summarise(min_date = min(time_rnd),
            max_date = max(time_rnd)
            )


actual_vs_pred_station10_line_graph <-
  all_models_preds$`10`[[10]] %>%
  filter(!is.na(keras) &
           !is.na(prophet) &
           !is.na(h2o) &
           !is.na(lag_05)
         ) %>% 
  arrange(station_id,
          time_rnd
          ) %>% 
  gather(key = "model",
         value = "bikes_avail_now",
         actual,
         keras,
         # prophet, # removing prophet because the prediction is so poor it's not worth ploting
         h2o,
         lag_05
         ) %>% 
  mutate(line_width = case_when(model == "actual" ~ 0.30,
                                TRUE ~ 0.15
                                )
         ) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now,
             colour = model,
             linetype = model
             )
         ) +
  geom_line() +
  coord_cartesian(xlim = c(as_datetime("2014-11-11 12:00:00"),
                           as_datetime("2014-11-11 18:00:00")
                           ),
                  ylim = c(0, 8)
                  ) +
  scale_x_datetime(breaks = c(as_datetime("2014-11-11 12:00:00"),
                              as_datetime("2014-11-11 14:00:00"),
                              as_datetime("2014-11-11 16:00:00"),
                              as_datetime("2014-11-11 18:00:00")
                              ),
                   date_labels = "%H:%M"
                   ) +
  scale_y_continuous(breaks = seq(0, 8, 2)) +
  scale_color_manual(values = c("black", "red", "blue", "green")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(title = "Bike Availability Comparison",
       subtitle = "Station ID: 10",
       x = "Time (2014-11-11, Tue)",
       y = "# Bikes Available"
       ) +
  NULL

# actual_vs_pred_station10_line_graph + theme_minimal()
# actual_vs_pred_station10_line_graph + theme_bw()
actual_vs_pred_station10_line_graph


ggsave(filename = "actual_vs_pred_station10_line_graph.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```
  
    
  Plot an example of the actual vs. predictions over time for `station_id` = 61.
```{r}

actual_vs_pred_station61_line_graph <-
  all_models_preds$`61`[[10]] %>%
  filter(!is.na(keras) &
           !is.na(prophet) &
           !is.na(h2o) &
           !is.na(lag_05)
         ) %>% 
  arrange(station_id,
          time_rnd
          ) %>% 
  gather(key = "model",
         value = "bikes_avail_now",
         actual,
         keras,
         prophet, # removing prophet because the prediction is so poor it's not worth ploting
         h2o,
         lag_05
         ) %>% 
  mutate(line_width = case_when(model == "actual" ~ 0.30,
                                TRUE ~ 0.15
                                )
         ) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now,
             colour = model,
             linetype = model
             )
         ) +
  geom_line() +
  coord_cartesian(xlim = c(as_datetime("2014-11-18 06:00:00",
                                       tz = "America/Los_Angeles"
                                       ),
                           as_datetime("2014-11-18 21:00:00",
                                       tz = "America/Los_Angeles"
                                       )
                           )
                  ) +
  scale_x_datetime(date_breaks = "2 hours",
                   timezone = "America/Los_Angeles",
                   date_labels = "%H:%M"
                   ) +
  scale_color_manual(values = c("black", "red", "blue", "green", "purple")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(title = "Bike Availability Comparison",
       subtitle = "Station ID: 61  (15 hours on Tue, 2014-11-18)",
       x = "Time",
       y = "# Bikes Available"
       ) +
  NULL

# actual_vs_pred_station10_line_graph + theme_minimal()
# actual_vs_pred_station10_line_graph + theme_bw()
actual_vs_pred_station61_line_graph


ggsave(filename = "actual_vs_pred_station61_line_graph.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```
  
    
  Plot an example of the actual vs. predictions over time for `station_id` = 61, then save the result as a .gif (ANIMATED!) file.
```{r}

actual_vs_pred_station61_line_graph_anim <-
  all_models_preds$`61`[[10]] %>%
  filter(!is.na(keras) &
           !is.na(prophet) &
           !is.na(h2o) &
           !is.na(lag_05) &
           between(x = time_rnd,
                   lower = as_datetime("2014-11-18 06:00:00",
                                       tz = "America/Los_Angeles"
                                       ),
                   upper = as_datetime("2014-11-18 21:00:00",
                                       tz = "America/Los_Angeles"
                                       ),
                   incbounds = TRUE
                   )
         ) %>% 
  arrange(station_id,
          time_rnd
          ) %>% 
  gather(key = "model",
         value = "bikes_avail_now",
         actual,
         keras,
         prophet, # removing prophet because the prediction is so poor it's not worth ploting
         h2o,
         lag_05
         ) %>% 
  mutate(line_width = case_when(model == "actual" ~ 0.30,
                                TRUE ~ 0.15
                                )
         ) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now,
             colour = model,
             linetype = model,
             group = model
             )
         ) +
  geom_line() +
  scale_x_datetime(date_breaks = "2 hours",
                   timezone = "America/Los_Angeles",
                   date_labels = "%H:%M"
                   ) +
  scale_color_manual(values = c("black", "red", "blue", "green", "purple")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(title = "Bike Availability Comparison",
       subtitle = "Station ID: 61  (15 hours on Tue, 2014-11-18)",
       x = "Time",
       y = "# Bikes Available"
       ) +
  NULL


actual_vs_pred_station61_line_graph_anim

actual_vs_pred_station61_line_graph_anim +
  transition_reveal(time_rnd) +
  # ease_aes("linear") +
  ease_aes("cubic-in-out")


anim_save(filename = "actual_vs_pred_station61_line_graph_anim.gif",
          path = paste0(wd,
                        "/Viz/"
                        )
          )


```
  
    
  Plot an example of the actual vs. predictions over time for `station_id` = 61, then save the result as a .mpeg (VIDEO!) file.
```{r}

actual_vs_pred_station61_line_graph_mpeg <-
  all_models_preds$`61`[[10]] %>%
  filter(!is.na(keras) &
           !is.na(prophet) &
           !is.na(h2o) &
           !is.na(lag_05) &
           between(x = time_rnd,
                   lower = as_datetime("2014-11-18 06:00:00",
                                       tz = "America/Los_Angeles"
                                       ),
                   upper = as_datetime("2014-11-18 21:00:00",
                                       tz = "America/Los_Angeles"
                                       ),
                   incbounds = TRUE
                   )
         ) %>% 
  arrange(station_id,
          time_rnd
          ) %>% 
  gather(key = "model",
         value = "bikes_avail_now",
         actual,
         keras,
         prophet, # removing prophet because the prediction is so poor it's not worth ploting
         h2o,
         lag_05
         ) %>% 
  mutate(line_width = case_when(model == "actual" ~ 0.30,
                                TRUE ~ 0.15
                                )
         ) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now,
             colour = model,
             linetype = model,
             group = model
             )
         ) +
  geom_line() +
  scale_x_datetime(date_breaks = "2 hours",
                   timezone = "America/Los_Angeles",
                   date_labels = "%H:%M"
                   ) +
  scale_color_manual(values = c("black", "red", "blue", "green", "purple")) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(title = "Bike Availability Comparison",
       subtitle = "Station ID: 61  (15 hours on Tue, 2014-11-18)",
       x = "Time",
       y = "# Bikes Available"
       ) +
  NULL


actual_vs_pred_station61_line_graph_mpeg


actual_vs_pred_station61_line_graph_mpeg <-
  actual_vs_pred_station61_line_graph_mpeg +
  transition_reveal(time_rnd) +
  ease_aes("linear")
  

animate(actual_vs_pred_station61_line_graph_mpeg,
        renderer = av_renderer()
        )


anim_save(filename = "actual_vs_pred_station61_line_graph_mpeg.mpeg",
          path = paste0(wd,
                        "/Viz/"
                        )
          )

```


