---
title: "SF Bikeshare Status - Viz for Presentation to Transportation Techies Bike Hack Night X"
output: html_notebook
---


**NOTE:** before running this chunk, the AWS Instance Type was switched from t2.micro to t3.2xlarge.

**NOTE:** `model_setup.Rmd` should be run first to create the needed datasets for creation of the viz.



## Setup  
  Load the relevant libraries.
```{r}
# {r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


library("tidyverse")
library("data.table")
library("lubridate")
library("ggmap")
library("gganimate")
library("gifski")
library("transformr")
library("av")


library("DBI")
# install.packages("RSQLite")
library("RSQLite")

```
    
  Session Info.
```{r}

sessionInfo()

```
  
    
  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

# Use below when running from AWS
# opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/Bikeshare_Status_SF/")

# Use below when running from laptop
opts_knit$set(root.dir = "/Users/mdturse/Dropbox/_AWS/Bikeshare_Status_SF/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
    
    
  Remove unneeded files.
```{r}

rm(lmtd_vars_lag3)

```


## Create Line Graphs of Bike Availability Over Time
    
  Individual line graph for each station - uses all available data (aproximately 2 years).
```{r}

dates <-
  small_test %>% 
  map(~summarise(.x,
                 date_min = as_date(min(time_rnd, na.rm = TRUE)),
                 date_max = as_date(max(time_rnd, na.rm = TRUE))
                 ) %>% 
        mutate(days = as.numeric(date_max - date_min),
               years = format(round(days / 365.15,
                                    digits = 1
                                    ),
                              nsmall = 1
                              )
               ) %>% 
        pull(years)
      )

# dates$`10`


all_dates <-
  pmap(.l = list(a = small_test,
               b = names(small_test),
               c = dates
               ),
     .f = function(a, b, c) {
       data = a %>% select(time_rnd, bikes_avail_now)
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line(color = "blue") +
         theme_bw() +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (", c, " years of data)"),
              x = "Date",
              y = "# Bikes Available"
              )
       }
     )


all_dates


pmap(.l = list(a = all_dates,
               b = names(all_dates)
               ),
     .f = function(a, b) {
       ggsave(plot = a,
              filename = paste0("bikes_avail_01_", b, "_all_dates.png"),
              path = paste0(wd,
                            "/Viz/"
                            )
              )
       }
     )

```
  
    
  Faceted line graph for each station - uses all available data (aproximately 2 years).
```{r}

all_dates_facet <-
  small_test %>% 
  bind_rows() %>% 
  select(station_id, time_rnd, bikes_avail_now) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now
                    )
         ) +
  geom_line() +
  facet_wrap(~station_id) +
theme_bw() +
  labs(title = "Bike Availability",
       subtitle = "By Station ID",
       x = "Date",
       y = "# Bikes Available"
       )



all_dates_facet


ggsave(plot = all_dates_facet,
       filename = "all_dates_facet.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```
  
  Faceted line graph for each station, colored by station - uses all available data (aproximately 2 years). **NOTE:** It takes considerable time to create the .gif file because of the amount of data involved.
```{r}

all_dates_facet_color <-
  small_test %>% 
  bind_rows() %>% 
  select(station_id, time_rnd, bikes_avail_now) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now,
             colour = station_id
             )
         ) +
  geom_line() +
  facet_wrap(~station_id) +
theme_bw() +
  labs(title = "Bike Availability",
       subtitle = "By Station ID  (2 years of data)",
       x = "Date",
       y = "# Bikes Available"
       ) +
  theme(legend.position = "none") +
  NULL



all_dates_facet_color

ggsave(plot = all_dates_facet_color,
       filename = "all_dates_facet_color.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )


# Also save a .gif version
all_dates_facet_color +
  transition_reveal(time_rnd) +
  ease_aes("linear")

anim_save(filename = "all_dates_facet_color.gif",
          path = paste0(wd,
                        "/Viz/"
                        )
          )

```
  
    
  Line graphs for each station, colored by station, overlayed in a single graph - uses all available data (aproximately 2 years). **NOTE:** This looks messy and is not worth saving.
```{r}

all_dates_combo <-
  small_test %>% 
  bind_rows() %>% 
  select(station_id, time_rnd, bikes_avail_now) %>% 
  ggplot(aes(x = time_rnd,
             y = bikes_avail_now,
             colour = station_id
             )
         ) +
  geom_line() +
  theme_bw() +
  labs(title = "Bike Availability",
       subtitle = "By Station ID",
       x = "Date",
       y = "# Bikes Available"
       ) +
  theme(legend.position = "none") +
  NULL



all_dates_combo


# ggsave(plot = all_dates_combo,
#        filename = "all_dates_combo.png",
#        path = paste0(wd,
#                      "/Viz/"
#                      )
#        )

```
  
    
  Graph one week of data - one graph per station.
```{r}

one_week <-
  pmap(.l = list(a = small_test,
               b = names(small_test)
               ),
     .f = function(a, b) {
       data = a %>% select(time_rnd, bikes_avail_now)
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line(color = "blue") +
         coord_cartesian(xlim = c(as_datetime("2014-11-17 00:00:00", # Monday
                                              tz = "America/Los_Angeles"
                                              ),
                                  as_datetime("2014-11-24 00:00:00", # Monday
                                              tz = "America/Los_Angeles"
                                              )
                                  )
                         ) +
         theme_bw() +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (1 week of data)"),
              x = "Date",
              y = "# Bikes Available"
              )
       }
     )


one_week


pmap(.l = list(a = one_week,
               b = names(one_week)
               ),
     .f = function(a, b) {
       ggsave(plot = a,
              filename = paste0("bikes_avail_02_", b, "_1_week.png"),
              path = paste0(wd,
                            "/Viz/"
                            )
              )
       }
     )

```
  
    
  Animate one week of data, just for `station_id` = 61.
```{r}

one_week_anim <-
  pmap(.l = list(a = small_test,
               b = names(small_test)
               ),
     .f = function(a, b) {
       data = a %>% 
         select(time_rnd, bikes_avail_now) %>% 
         filter(between(x = time_rnd,
                        lower = as_datetime("2014-11-17 00:00:00", # Monday
                                            tz = "America/Los_Angeles"
                                            ),
                        upper = as_datetime("2014-11-24 00:00:00", # Monday
                                            tz = "America/Los_Angeles"
                                            ),
                        incbounds = TRUE
                        )
                )
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line() +
         coord_cartesian(ylim = c(0, 30)) +
         scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30)
                            ) +
         
         theme_bw() +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (1 week of data)"),
              x = "Date",
              y = "# Bikes Available"
              )
       }
     )


one_week_anim

one_week_anim$`61` +
  transition_reveal(time_rnd) +
  ease_aes("linear")

anim_save(filename = "anim_one_week_61.gif",
          path = paste0(wd,
                        "/Viz/"
                        )
          )

```
  
    
  Animate one week of data, faceted and colored by each `station_id` value.
```{r}

one_week_facet_anim <-
  small_test %>% 
  bind_rows() %>% 
  select(station_id, time_rnd, bikes_avail_now) %>% 
  filter(between(x = time_rnd,
                 lower = as_datetime("2014-11-17 00:00:00", # Monday
                                     tz = "America/Los_Angeles"
                                     ),
                 upper = as_datetime("2014-11-24 00:00:00", # Monday
                                     tz = "America/Los_Angeles"
                                     ),
                 incbounds = TRUE
                 )
         ) %>% 
  ggplot(aes(x = time_rnd,
           y = bikes_avail_now,
           colour = station_id
           )
                ) +
  geom_line() +
  facet_wrap(~station_id) +
  theme_bw() +
         labs(title = "Bike Availability",
              subtitle = "By Station ID  (1 week of data)",
              x = "Date",
              y = "# Bikes Available"
              ) +
  theme(legend.position = "none")


one_week_facet_anim

one_week_facet_anim +
  transition_reveal(time_rnd) +
  ease_aes("linear")


anim_save(filename = "one_week_facet_anim.gif",
          path = paste0(wd,
                        "/Viz/"
                        )
          )

```
  
    
  Graph two days of data, one graph for each `station_id` value.
```{r}

two_days <-
  pmap(.l = list(a = small_test,
               b = names(small_test)
               ),
     .f = function(a, b) {
       data = a %>% select(time_rnd, bikes_avail_now)
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line(color = "blue") +
         # All Tue & all Wed
         coord_cartesian(xlim = c(as_datetime("2014-11-18 00:00:00",
                                              tz = "America/Los_Angeles"
                                              ),
                                  as_datetime("2014-11-20 00:00:00",
                                              tz = "America/Los_Angeles"
                                              )
                                  )
                         ) +
         theme_bw() +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (2 days of data)"),
              x = "Date & Time",
              y = "# Bikes Available"
              )
       }
     )


two_days


pmap(.l = list(a = two_days,
               b = names(two_days)
               ),
     .f = function(a, b) {
       ggsave(plot = a,
              filename = paste0("bikes_avail_03_", b, "_2_days.png"),
              path = paste0(wd,
                            "/Viz/"
                            )
              )
       }
     )

```
  
    
  Animate two days of data, just for `station_id` = 61.
```{r}

two_days_anim <-
  pmap(.l = list(a = small_test,
               b = names(small_test)
               ),
     .f = function(a, b) {
       data = a %>% 
         select(station_id, time_rnd, bikes_avail_now) %>% 
         # All Tue and all Wed
         filter(between(x = time_rnd,
                        lower = as_datetime("2014-11-18 00:00:00",
                                              tz = "America/Los_Angeles"
                                              ),
                        upper = as_datetime("2014-11-20 00:00:00",
                                              tz = "America/Los_Angeles"
                                              ),
                        incbounds = TRUE
                        )
                )
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line(color = "blue") +
         theme_bw() +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (2 days of data)"),
              x = "Date & Time",
              y = "# Bikes Available"
              ) +
         theme(legend.position = "none") +
         NULL
       }
     )


two_days_anim

two_days_61_anim <-
  two_days_anim$`61` +
  transition_reveal(time_rnd) +
  ease_aes("linear")

two_days_61_anim


# animate as a .gif
anim_save(filename = "two_days_61_anim.gif",
          path = paste0(wd,
                        "/Viz/"
                        )
          )
  

# animate as a .mpeg
animate(two_days_61_anim,
        renderer = av_renderer()
        )


anim_save(filename = "two_days_61_anim.mpeg",
          path = paste0(wd,
                        "/Viz/"
                        )
          )

```
  
    
  Graph one day of data, one graph for each `station_id` value.
```{r}

one_day <-
  pmap(.l = list(a = small_test,
               b = names(small_test)
               ),
     .f = function(a, b) {
       data = a %>% select(time_rnd, bikes_avail_now)
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line(color = "blue") +
         # All Tue
         coord_cartesian(xlim = c(as_datetime("2014-11-18 00:00:00",
                                              tz = "America/Los_Angeles"
                                              ),
                                  as_datetime("2014-11-19 00:00:00",
                                              tz = "America/Los_Angeles"
                                              )
                                  )
                         ) +
         theme_bw() +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (1 day of data)"),
              x = "Date & Time",
              y = "# Bikes Available"
              )
       }
     )


one_day


pmap(.l = list(a = one_day,
               b = names(one_day)
               ),
     .f = function(a, b) {
       ggsave(plot = a,
              filename = paste0("bikes_avail_04_", b, "_1_day.png"),
              path = paste0(wd,
                            "/Viz/"
                            )
              )
       }
     )

```
  
    
  Graph 15 hours of data, one graph per `station_id` value.
```{r}

fifteen_hours <-
  pmap(.l = list(a = small_test,
               b = names(small_test)
               ),
     .f = function(a, b) {
       data = a %>% select(time_rnd, bikes_avail_now)
       
       data %>% 
         ggplot(aes(x = time_rnd,
                    y = bikes_avail_now
                    )
                ) +
         geom_line(color = "blue") +
         theme_bw() +
         # 15 hours on Tue
         coord_cartesian(xlim = c(as_datetime("2014-11-18 06:00:00",
                                              tz = "America/Los_Angeles"
                                              ),
                                  as_datetime("2014-11-18 21:00:00",
                                              tz = "America/Los_Angeles"
                                              )
                                  )
                         ) +
         labs(title = "Bike Availability",
              subtitle = paste0("Station ID: ", b, "  (15 hours of data)"),
              x = "Date & Time",
              y = "# Bikes Available"
              )
       }
     )


fifteen_hours


pmap(.l = list(a = fifteen_hours,
               b = names(fifteen_hours)
               ),
     .f = function(a, b) {
       ggsave(plot = a,
              filename = paste0("bikes_avail_05_", b, "_15_hours.png"),
              path = paste0(wd,
                            "/Viz/"
                            )
              )
       }
     )


```


## Create Maps of Station Locations  
    
  Get the data from the SqlLite database.
```{r}

# connect to the sqlite database
con <- dbConnect(SQLite(),
                 paste0(wd,
                        "/Data/Raw/",
                        "database.sqlite"
                        )
                 )

# get a list of the talbles
as.data.frame(dbListTables(con))

# get data for every 5 minutes
myQuery <- dbSendQuery(con,
                       "Select          *
                        From            status
                          Where         (     time like '%:_0:__'
                                          or  time like '%:_5:__'
                                         )
                       "
                       )

sf_status <-
  dbFetch(myQuery, n = -1) %>% 
  as.data.table

class(sf_status)
glimpse(sf_status)
object.size(sf_status)
summary(sf_status)
sf_status %>% select(station_id) %>% distinct() %>% nrow()

dbClearResult(myQuery)

# View(sf_status %>% head(500))
```
  
    
  For testing purposes, select only six `station_id` values - the two most frequently used stations, the two least frequently used stations, and two in the middle.
```{r}

avg_bikes_available <-
  sf_status[ ,
            .(bike_avail_cnt = .N,
              bike_avail_avg = mean(bikes_available, na.rm = TRUE)
              ),
            by = station_id
            ][order(bike_avail_cnt,
                    bike_avail_avg
                    )
              ][ , row_num := row_number(bike_avail_cnt)]

# View(avg_bikes_available)

station_id_2top <- # 22 & 61
  avg_bikes_available %>% 
  top_n(n = 2,
        wt = row_num
        ) %>% 
  pull(station_id)

station_id_2bot <- # 84 & 83
  avg_bikes_available %>% 
  top_n(n = -2,
        wt = row_num
        ) %>% 
  pull(station_id)

station_id_2mid <- # 10 & 21
  avg_bikes_available %>% 
  filter(row_num == nrow(avg_bikes_available) / 2 |
           row_num == (nrow(avg_bikes_available) / 2) + 1
         ) %>% 
  pull(station_id)

station_id_select <-
  c(station_id_2top, station_id_2bot, station_id_2mid)

station_id_select


rm(station_id_2top, station_id_2bot, station_id_2mid, sf_status, station_id_select)

```
  
    
  Get the data from the SqlLite database for stations.
```{r}

# connect to the sqlite database
con <- dbConnect(SQLite(),
                 paste0(wd,
                        "/Data/Raw/",
                        "database.sqlite"
                        )
                 )

# get a list of the talbles
as.data.frame(dbListTables(con))

# get data for stations
myQuery <- dbSendQuery(con,
                       "Select        *
                        From          station
                        Where         id in (22, 61, 84, 83, 10, 21)
                       "
                       )

sf_stations <-
  dbFetch(myQuery, n = -1) %>% 
  mutate(avg_bikes_avail_group = case_when(id == 22 | id == 61 ~ "most",
                                           id == 84 | id == 83 ~ "least",
                                           id == 10| id == 21 ~ "middle"
                                           ),
         station_and_group = paste0(id, "_", avg_bikes_avail_group)
         )

class(sf_stations)
glimpse(sf_stations)
object.size(sf_stations)
summary(sf_stations)

dbClearResult(myQuery)

# View(sf_stations %>% arrange(lat, long))


rm(con, myQuery)

```
  
    
  Map all six `station_id` values.
```{r}

stations <- sf_stations %>% 
  select(long, lat)


register_google(key = "--ADD YOUR GOOGLE API KEY HERE--")

map <-
  get_googlemap(center = c(long = -122.185141, lat = 37.562057), # SF Bay
                zoom = 10,
                maptype = "roadmap"
                )

map_all_stations <-
  ggmap(map) +
  geom_label(data = sf_stations,
            aes(x = long,
                y = lat,
                label = station_and_group,
                fill = avg_bikes_avail_group
                ),
            colour = "white",
            fontface = "bold"
            ) +
  labs(title = "Postion of Analyzed Stations",
       subtitle = "Entire Bay Area",
       x = "Longitude",
       y = "Latitude"
       ) +
  theme(legend.position = "none") +
  NULL


map_all_stations


ggsave(filename = "map_all_stations.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )


```
  
    
  Map just San Francisco stations:  1 most.
```{r warning=FALSE}

sf <-
  sf_stations %>% 
  filter(id == 61)


map <-
  get_googlemap(center = c(long = sf$long, lat = sf$lat), # SF Bay
                zoom = 15,
                maptype = "roadmap"
                )

map_sf <-
  ggmap(map) +
  geom_label(data = sf_stations,
            aes(x = long,
                y = lat,
                label = station_and_group,
                fill = avg_bikes_avail_group
                ),
            colour = "white",
            fontface = "bold"
            ) +
  theme(legend.position = "none") +
  labs(title = "Postion of Analyzed Stations",
       subtitle = "San Francisco",
       x = "Longitude",
       y = "Latitude"
       ) +
  theme(legend.position = "none") +
  NULL


map_sf


ggsave(filename = "map_sf.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```
  
    
  Map just Redwood City: 1 most, 1 middle, 1 least.
```{r warning=FALSE}

rc <-
  sf_stations %>% 
  filter(id == 22)


map <-
  get_googlemap(center = c(long = rc$long, lat = rc$lat), # SF Bay
                zoom = 15,
                maptype = "roadmap"
                )

map_rc <-
  ggmap(map) +
  geom_label(data = sf_stations,
            aes(x = long,
                y = lat,
                label = station_and_group,
                fill = avg_bikes_avail_group
                ),
            colour = "white",
            fontface = "bold"
            ) +
  theme(legend.position = "none") +
  labs(title = "Postion of Analyzed Stations",
       subtitle = "Redwood City",
       x = "Longitude",
       y = "Latitude"
       ) +
  theme(legend.position = "none") +
  NULL


map_rc


ggsave(filename = "map_rc.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```
  
    
  Map just San Jose: 1 middle & 1 least.
```{r warning=FALSE}

sj <-
  sf_stations %>% 
  filter(id == 10)


map <-
  get_googlemap(center = c(long = sj$long, lat = sj$lat), # SF Bay
                zoom = 15,
                maptype = "roadmap"
                )

map_sj <-
  ggmap(map) +
  geom_label(data = sf_stations,
            aes(x = long,
                y = lat,
                label = station_and_group,
                fill = avg_bikes_avail_group
                ),
            colour = "white",
            fontface = "bold"
            ) +
  theme(legend.position = "none") +
  labs(title = "Postion of Analyzed Stations",
       subtitle = "San Jose",
       x = "Longitude",
       y = "Latitude"
       ) +
  theme(legend.position = "none") +
  NULL


map_sj


ggsave(filename = "map_sj.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```


## Create a Cleaner Plot of Variable Importance From H2O (for Statin_Id = 61)  
  
  Load up H2O files.
```{r}

automl_models_h2o <-
  read_rds(path = paste0(wd,
                         "/Models/",
                         "automl_models_h2o.rds"
                         )
           )

names(automl_models_h2o)
library("h2o")
h2o.init()

```
  
    
  H2O Created Variable Importance Plot, for `station_id` = 61.
```{r}

leader_models_varimp <-
  pmap(.l = list(a = automl_models_h2o),
       .f = function(a) {
         h2o.varimp(object = a@leader)
         }
       )

# leader_models_varimp_plot <-
#   pmap(.l = list(a = automl_models_h2o),
#        .f = function(a) {
#          h2o.varimp_plot(model = a@leader, num_of_features = 10)
#          }
#        )

leader_models_varimp$`61`
h2o.varimp_plot(model = automl_models_h2o$`61`@leader, num_of_features = 10)

```
  
    
  Manually produced variable importance plot for `station_id` = 61.
```{r}

leader_models_varimp$`61` %>% 
  as.data.frame() %>% 
  head(10) %>% 
  arrange(desc(variable)) %>% 
  ggplot(aes(x = reorder(variable, scaled_importance),
             y = scaled_importance
             )
         ) +
    geom_col(fill = "blue") +
  geom_text(aes(label = format(round(scaled_importance,
                                     digits = 2
                                     ),
                               nsmall = 2
                               )
                ),
            hjust = -0.05,
            size = 3
            ) +
    coord_flip() +
    theme_bw() +
  labs(title = "Variable Importance (GBM)",
       subtitle = "Station ID:  61",
       x = "Variable",
       y = "Scaled Importance"
       ) +
    NULL
  

ggsave(filename = "var_imp_plot_61.png",
       path = paste0(wd,
                     "/Viz/"
                     )
       )

```

